---
title       : Creating slideshows, latex, and functions
subtitle    : 
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
ext_widgets: {rCharts: [libraries/nvd3]}

---
## Slidify

1. Easy to use software that links R, markdown, and html5 slide decks

```
library(devtools)
install_github('slidify', 'ramnathv')
```

---
## To get started:
```
library(slidify)
author("mydeck")
``` 

This will do the following:

1. Create a new directory named mydeck.
2. Copy necessary scaffolding for the presentation.
3.  Open index.Rmd for you to edit.


---
```
---
title       : 
subtitle    : 
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---
```

---
## Getting started

* Seperate slides by an empty line followed by three dashes (---)
* Give your slides a header, and use markdown within the slide:

``` 
---
 ## This is the header for this slides
  * These are bullets
    + So is this
```


---
## Inputting code and graphics

```
{r,echo=F,fig.height=8,warning=F,message=F}
library(ggplot2)
ggplot(mtcars,aes(x=mpg,y=hp))+geom_smooth(method="lm")
```

---
## Inputting code and graphics
```{r,echo=F,fig.height=8,warning=F,message=F}
library(ggplot2)
ggplot(mtcars,aes(x=mpg,y=hp))+geom_smooth(method="lm")
```

---
## Can input web pages with iframe
```
<iframe src = 'https://github.com/ylelkes/R_wav' height='600px'></iframe>
```


--- 

<iframe src = 'http://www.uva.nl/en/home' height='600px'></iframe>

---
## Interactive Chart

```{r echo = F, results = 'asis',comment=NA}
library(rCharts)
hair_eye_male = subset(as.data.frame(HairEyeColor), Sex == "Male")
m1 <- nPlot(Freq ~ Hair, group = 'Eye', 
      data = hair_eye_male, type = 'multiBarChart'
    )
m1$set(pointSize = 0, lineWidth = 1)
m1$print('chart2', include_assets = TRUE)
```

---
The following frameworks are compatible with Slidify for making your presentations:

- [io2012](https://code.google.com/p/io-2012-slides/)
- [html5slides](https://code.google.com/p/html5slides/)
- [deck.js](http://imakewebthings.com/deck.js/)
- [dzslides](http://paulrouget.com/dzslides/)
- [landslide](https://github.com/adamzap/landslide)
- [Slidy](http://www.w3.org/Talks/Tools/Slidy2/Overview.html#)

---
## Exercise
* Create a 4 slide presentation:
  + Title Slide
  + Some Bullet points and header
  + A plot
  + A proportion table (try pander::pander())
  
* Try different frameworks

---
```{r pander}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)
library(pander)
panderOptions("digits", 2)
pander(out)
```

---
## Latex/Knitr
* Latex is a popular typesetting program. 
* Advantages:
  + Looks good
  + You don't have to worry about headings, sections, etc
  + Combine it with knitr and you have some reproducible research

---
## Getting Started

```
% hello.tex - Our first LaTeX example!

\documentclass{article}

\begin{document}

Hello World!

\end{document}
```

---
```
\documentclass{article}
\usepackage{authblk}
\title{Introduction to Latex}
\author{Yphtach Lelkes}
\affil{ASCoR}
\date{\today}
\begin{document}
\SweaveOpts{concordance=TRUE}

\maketitle
\begin{abstract}
Put abstract here
\end{abstract}

Main text...
\end{document}
```

---
* With latex, we don't have to worry about section heads

\section{Section Name}

Text

\subsection{Subsection Name}

Text

\subsubsection{SubSubsection Name}

Text

\paragraph{Paragraph Name}

Text

---
* We also have an easy way to generate and build bibliographies.
* Different versions but we'll start with natbib
* google scholar example
* \citep{}
* \citet{}
* \cite{}
* Add \addbibresource{bibliography.bib} to preamble
* \printbibliography Where you want the bibliography to be


---
* Can include graphics:
  
  \begin{figure}
  \caption{Figure 1}
  \includegraphics[width=2.5cm, height=2.5]{img.xxx}
  \label{fig:figa}
  \end{figure}
+ \includegraphics[width=\textwidth]{img.xxx}

---
* Just like markdown, knitr/sweave uses chunks

> <<>>=

> CODE

> @

* see http://yihui.name/knitr/options/

---
* If we want to include figures, we can do something like:

> <<plot1, fig.pos="t", fig.height=4, fig.width=4, fig.cap="First plot",echo=F>>=

> plot(mtcars$gear~mtcars$cyl)

> @

* We can later refer to that figure, e.g., See Figure \ref{fig:plot1}, which will look something like: See Figure 1. 

---
* If we want to include tables, we can do something like:

> <<results=tex, echo=F>>=
>
> print(xtable(prop.table(table(mtcars$cyl,mtcars$gear)),caption="This is table1",  label="tab:t1"))
>
> @

* We can later refer to that table, e.g., See Table \ref{tab:t1}, which will look something like: See Table 1. 
* Just like inline code in markdown, R uses the term \Sexpr{6+6}


---
## Let's try it.

```
\documentclass[man]{apa6} % man for manuscript format, jou for journal format, doc for standard LaTeX document format
\usepackage[natbibapa]{apacite} % Divine intervention help you if you need to use a different citation package.
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}

\usepackage{url}   % this allows us to cite URLs in the text
\usepackage{graphicx}   % allows for graphic to float when doing jou or doc style
\usepackage{verbatim}   % allows us to use \begin{comment} environment
\usepackage{caption}
%\usepackage{lscape}
\usepackage{pdflscape}
```



---
* Add title information

```
\title{Your APA6-Style Manuscript}
\shorttitle{Your APA6-Style Manuscript}
\author{You}
\affiliation{Somewhere}

\abstract{Your abstract here.}
\begin{document}
```

---
## Put this where you want your bibliography

> 
> \bibliographystyle{apacite}
>
> \bibliography{bibexample} % NAME OF your bibliography file


---
* Make a short fake document with different sections and subsections.
* Use the mtcars data
* Include a graph and in the chunk and give the figure a title and a label
* Include a series of regressions and the stargazer package to include a regression table
* Give the table a caption and a label (see xtable options). 
* Refer to the table in the text with \ref{}
* Create a chunk where you define x <- 36 (or something like that). Use \Sexpr{x} in a sentence. 
* use \includegraphics to include the image hipdog.jpg and refer to that figure in the text with \ref{}
* Try the different citation options and generate a bibliography with bibexample.bib
* In the preamble see what happens when you change man to jou to doc

---
## Writing functions
1. Sometimes we find ourselves rewriting code over and over again. 
2. We can use functions to make plyr and other packages do what we want. 

---
##  for instance:
```{r}
square.it <- function(x) {
    square <- x * x
    return(square)
}
```

---
## 
  
```{r}
square.it(6)
```

---

```{r}
xy <- function(x,y) {
    xy <- x * y
    return(xy)
}
```
---

  
```{r}
xy(6,4)
```

---

* One function I use all the time  recodes all the values in a vector to lie between 0 and 1, where 0 indicates the lowest score in that vector, 1 indicates the highest. To calculate this, we substract the minimum of the vector X from X and divide that score  by the maximimum score of X minus the minimum of X. Write a zero1 function and apply it to mtcars$disp and mtcars$cyl. 

---
## Exercises
* Dplyr works with user-defined functions. The input into a function need not be a vector or a number, but it can also be a dataframe. Remember that dplyr pushes an object forward. 

* With the ToothGrowth data, write a function that takes a dataset (x), and regresses length (len) on dose, and keeps the estimate and standard error for the dose coefficient.

* Stick that function into dplyr so that you get different output based on supplement type, and plot the coefficients and 95 percent confidence intervals.

---
## More exercises
1. Let's create a list of 10 dataframes with x and y in each:

```{r}
sim_list = replicate(n = 10,                  
expr = {data.frame(x = rnorm(50), y = rnorm(50), z=rnorm(10))}, simplify = F)
```
2. Write a function, that, given a dataframe regresses y on x and z (y=x+z).
3. Input that function and dataframe into lapply(), which performs a function on each element in a list. The result will be a list of regressions.  
4. Put the output into stargazer,screenreg, or another regression to table function, so you have a regression table that is 10 columns long.    