---
title       : Finishing up reshape, dplyr
subtitle    : 
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## reshape2 package
* Reshape2 is based on two functions
* melt takes wide-format data and melts it into long-format data.
* dcast takes long-format data and casts it into wide-format data.

---
## melting data
* Have a dataset that we want to "melt", so that all outcome measures are stacked

```{r}
data <- read.table('http://www.ats.ucla.edu/stat/r/faq/hsb2.csv', header=T, sep=",")
head(data,5)
```

---
## melting data
```{r}
library(reshape2)
melteddata <- melt(data,id.vars = 1:6)
head(melteddata,10)
```

---
## melting data
```{r,warning=F,message=FALSE}
library(ggplot2)
ggplot(melteddata,aes(x=prog,y=value,group=prog))+geom_boxplot()+facet_wrap(~variable)
```

---
## recasting data
```{r}
backtowide <- dcast(melteddata,formula = id+female+race+ses+schtyp+prog~variable)
```

---
## recasting data
```{r}
backtowide_byrace <- dcast(melteddata,formula = race~variable,fun.aggregate = mean)
backtowide_byrace
```

---
## recasting data
```{r}
backtowide_byraceses <- dcast(melteddata,formula = race+ses~variable,fun.aggregate = mean)
backtowide_byraceses
```

---
## Exercises 

datasets:EuStockMarkets is a dataset in wide format that lists the closing prices for 4 EU stock markets
  - Create a plot with 4 lines that tracks closing prices by date.

datasets::airquality is a dataset that tracks 4 variables of climate by month and day.
  - melt the data so that measurements are stacked
  - cast so that you get a mean monthly measurement by outcome measure.
  
---

---
## dplyr

```
install.packages("dplyr")
library(dplyr)
flights <- tbl_df(read.csv(url("http://cdn.rawgit.com/ylelkes/R_wav/master/data/flights.csv")))
```
1. Most useful package (for me) in R. 
2. Another St. Wickham creation. 
3. A tool for data exploration and transformation

---
## 5 verbs
* filter: keep rows matching criteria
* select: pick columns by name
* arrange: reorder rows 
* mutate: add new variables
* summarise: reduce variables to values

--- 
## What is our general m.o. if wanted rows 1,3, & 4?
```{r}
df <- data.frame(   color = c("blue", "black", "blue", "blue", "black"),    value = 1:5)
head(df)
```

--- 
## What if we just wanted the "color" column

--- 
## What if we just wanted the "color" column
```{r}
select(df,color)
```

--- 
## Or if we didn't want the "color" column

--- 
## Or if we didn't want the "color" column
```{r}
select(df,-color)
```

--- 
## Take a look at the handout
```{r,echo=T}
library(dplyr)
flights <- tbl_df(read.csv("~/Dropbox/R_wav/data/flights.csv"))
names(flights)
```

* What are some ways of selecting columns "dep_delay" and "arr_delay"

---
## Selecting _delay

```
select(flights, arr_delay, dep_delay)
select(flights, arr_delay:dep_delay)
select(flights, ends_with("delay"))
select(flights, contains("delay"))
```

---
## Arrange 

```{r}
arrange(df,value)
```

---
## Arrange 

```{r}
arrange(df,desc(value))
```

---
## Exercises 
* Order the flights by departure date and time.
* Which flights were most delayed?
* Which flights caught up the most time during the flight?

---
## Exercises
* arrange(flights, date, hour, minute) 
* arrange(flights, desc(dep_delay))
* arrange(flights, desc(arr_delay))
* arrange(flights, desc(dep_delay - arr_delay))


---
## Arrange 

```{r}
mutate(df,double= 2* value)
```

---
## Arrange 

```{r}
mutate(df,double= 2* value,quadruple=2*double)
```

---
## Exercises
* Compute speed in mph from time (in minutes) and distance (in miles).
* Which flight flew the fastest?
* Add a new variable that shows how much time was made up or lost in flight.

---
## Exercises
```
flights <- mutate(flights, speed = distance / (air_time/60))
fastestflights <- arrange(flights, desc(speed))
flights <- mutate(flights, delta = dep_delay - arr_delay)
```

---
## Grouped Summaries
```{r}
summarise(df,total=sum(value))
```

---
## Grouped Summaries
```{r}
by_color <- group_by(df,color)
summarise(by_color,total=sum(value))
```

---
## Grouped summaries
* by_date <- group_by(flights, date)
* by_hour <- group_by(flights, date, hour)
* by_plane <- group_by(flights, plane)
* by_dest <- group_by(flights, dest) 

---
## Summary functions
* min(x), median(x), max(x), quantile(x, p)
* n(), n_distinct(), sum(x), mean(x) 
* sum(x > 10), mean(x > 10) 
* sd(x), var(x), iqr(x), mad(x)

---
## How might you summarise dep_delay for each day? 
- Include multiple summary statistics in the output

---
## Summarise by date

```
by_date <- group_by(flights, date)
delays <- summarise(by_date,   
mean = mean(dep_delay, na.rm = TRUE), 
median = median(dep_delay, na.rm = TRUE),
q75 = quantile(dep_delay, 0.75, na.rm = TRUE),
over_15 = mean(dep_delay > 15, na.rm = TRUE),
over_30 = mean(dep_delay > 30, na.rm = TRUE),
over_60 = mean(dep_delay > 60, na.rm = TRUE)  )
```

---
## Summarise by date
```
by_date <- group_by(flights, date)
no_missing <- filter(flights, !is.na(dep))
delays <- summarise(by_date,   
mean = mean(dep_delay), 
median = median(dep_delay),
q75 = quantile(dep_delay, 0.75),
over_15 = mean(dep_delay > 15),
over_30 = mean(dep_delay > 30),
over_60 = mean(dep_delay > 60)  )
```

---
## Chaining or Data Pipelines
Downside of functional interface is that it's  hard to read multiple operations:
```
hourly_delay <- filter(
  summarise(
  group_by(
  filter(
      flights,!is.na(dep_delay) ), date, hour ),  delay = mean(dep_delay), n = n() ), n > 10
)
```

---
## Dplyr allows for chaining
```
hourly_delay <- flights %>%
 filter(!is.na(dep_delay)) %>%
 group_by(date, hour) %>%
 summarise(delay = mean(dep_delay), n = n()) %>%
 filter(n > 10) 
```

---
## Exercises
Create data pipelines to answer the following
questions:

* Which destinations have the highest average
delays?
* Which flights (i.e. carrier + flight) happen every
day? Where do they fly to?
* On average, how do delays (of non-cancelled
flights) vary over the course of a day?
(Hint: hour + minute / 60)

---

flights %>%
 group_by(dest) %>%
 summarise(
 arr_delay = mean(arr_delay, na.rm = TRUE),
 n = n()) %>%
 arrange(desc(arr_delay))
 
---

flights %>%
 group_by(carrier, flight) %>%
 filter(n() == 365)
 
---
per_hour <- flights %>%
 filter(cancelled == 0) %>%
 mutate(time = hour + minute / 60) %>%
 group_by(time) %>%
 summarise(
 arr_delay = mean(arr_delay, na.rm = TRUE),
 n = n()
 ) 
 
 qplot(time, arr_delay, data = per_hour) 
