---
title: "Make these plots"
author: "ylelkes"
date: "September 7, 2015"
output: pdf_document
---

## Make these plots

Use the Fair data in the Ecdat package
[Read about it here](https://cran.r-project.org/web/packages/Ecdat/Ecdat.pdf)

## Part 1
```{r,echo=F,warning=F,message=F}
library(ggplot2)
library(Ecdat)
ggplot(Star,aes(tmathssk))+geom_histogram()
ggplot(Star,aes(tmathssk))+geom_density()
ggplot(Star,aes(tmathssk,fill=sex))+geom_density()
ggplot(Star,aes(tmathssk,fill=sex))+geom_density(alpha=.5)
ggplot(Star,aes(x=sex,y=tmathssk,fill=sex))+geom_boxplot(colour="purple")
ggplot(Star,aes(x=totexpk,y=tmathssk))+geom_smooth()
ggplot(Star,aes(x=totexpk,y=tmathssk))+geom_smooth()+geom_point()

library(dplyr)
names(Star)
Star$tmathssk
Star %>% group_by(sex,race) %>% summarise(math=mean(tmathssk,na.rm=T),read=mean(treadssk,na.rm=T),sd(tmathssk),N=length(treadssk)) -> ta

df <- data.frame(a = c(3:24), b = letters[1:22])
library(plotrix)
 twogrp<-c(rnorm(10)+4,rnorm(10)+20)
 gap.barplot(twogrp,gap=c(3,4),xlab="Index",ytics=c(3,6,17,20),
  ylab="Group values",main="Barplot with gap")

?gap.barplot
ggplot(mtcars,aes(mpg,disp))+geom_smooth(method="lm")+scale_x_continuous(breaks=seq(10,40,by=1),limits=c(0,40))+scale_y_continuous(limits=c(200,400))+ geom_hline(200,lty=1)
a <- lm($mpg~mtcars$disp)
b <- lm(mtcars$mpg~mtcars$hp)
stargazer::stargazer(a,b)
```

## Part 2
```{r,echo=F,warning=F,message=F}
ggplot(Star,aes(x=tmathssk,y=treadssk,colour=race))+geom_point()+facet_wrap(~sex)
ggplot(Star,aes(x=tmathssk,y=treadssk))+geom_point()+facet_grid(race~classk)
ggplot(Star,aes(x=tmathssk,y=treadssk,colour=sex))+geom_point()+facet_grid(race~classk)+geom_smooth(method=lm,colour="black")
ggplot(Star,aes(x=tmathssk,y=treadssk,shape=sex,colour=freelunk))+geom_point()+facet_grid(race~classk)+geom_smooth(method=lm)
```

Still have time? Change the axis labels, add a title

## Part 3
Here are two datasets:


```{r}
temperature_by_year <- 
  rio::import("https://rawgit.com/ylelkes/R_wav/master/data%20examples/temperature_by_year.RData")

futurepast <- 
  rio::import("https://rawgit.com/ylelkes/R_wav/master/data%20examples/climatetemperatures.RData")
```

The file [climatetemperatures.RData](https://github.com/ylelkes/R_wav/blob/master/data%20examples/climatetemperatures.RData?raw=true) . Temperature_by_year is used in the first plot; futurepast is based off of 15 or so climate models predicting the temperature of european countries in the end of the 21st centure. Hint: The 95 percent confidence intervals (which are plotted below are 1.96 times the standard error over/under the mean)

```{r,echo=F,warning=F,message=F,fig.width=8}
ggplot(temperature_by_year,aes(x=year,ymin=min,ymax=max,y=temperature))+geom_ribbon(alpha=.5,fill="red")+geom_line()
```

```{r,echo=F,warning=F,message=F,fig.height=8}
ggplot(futurepast,aes(x=temperature,xmin=temperature-1.96*se,xmax=temperature+1.96*se,y=locator,colour=time))+geom_point()+geom_errorbarh()
```


## Part 4 
Scalesand themes
```{r,echo=F}
drv_labels <- c("4" = "4wd", "f" = "fwd", "r" = "rwd")
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = drv)) +  
  scale_colour_manual(labels = drv_labels,values = c("purple","orange","grey","pink"))
```


Spend a few minutes reading about [themes](http://docs.ggplot2.org/current/theme.html)
Can you rotate the x-axis text and change the colour?
Can you change the colour and size of the y-axis?

```{r,echo=F}
drv_labels <- c("4" = "4wd", "f" = "fwd", "r" = "rwd")
ggplot(mpg, aes(manufacturer,hwy)) + 
  geom_point(aes(colour = drv)) +  
  scale_colour_manual(labels = drv_labels,values = c("purple","orange","grey","pink"))+theme(axis.text.x = element_text(angle = 90, hjust = 1,colour="black"))+theme(axis.text.y = element_text(size=15,colour="black"))
```
## RColor Brewer and ggthemr

* How close to this graph can you get?
* some hints: 
  * to order, try reorder(Origin, Total)
  * Use coord_flip()
  * You'll need RColorBrewer and ggthemes

```{r}
asylumdata <- 
  rio::import("https://rawgit.com/ylelkes/R_wav/master/data%20examples/asylumdata.csv")
```



```{r,echo=F,fig.height=8}
library("RColorBrewer")
library(ggthemes)
colourCount = length(unique(asylumdata$Origin))
getPalette = colorRampPalette(brewer.pal(3, "YlGnBu"))
ggplot(asylumdata, aes(reorder(Origin, Total),Total,fill=(reorder(Origin, Total))))+geom_bar(stat="identity")+coord_flip()+ylab("Number of Asylum Seekers")+xlab("Country of Origin")+theme_economist()+theme(legend.position="none")+scale_fill_manual(values = getPalette(colourCount))+ggtitle("Number of Refugees to Europe by Country")
```
