---
title       : Finishing up reshape, dplyr
subtitle    : 
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## reshape2 package
* Reshape2 is based on two functions
* melt takes wide-format data and melts it into long-format data.
* dcast takes long-format data and casts it into wide-format data.

---
## melting data
* Have a dataset that we want to "melt", so that all outcome measures are stacked

```{r}
data <- read.table('http://www.ats.ucla.edu/stat/r/faq/hsb2.csv', header=T, sep=",")
head(data,5)
```

---
## melting data
```{r}
library(reshape2)
melteddata <- melt(data,id.vars = 1:6)
head(melteddata,10)
```

---
## melting data
```{r}
library(ggplot2)
ggplot(melteddata,aes(x=prog,y=value,group=prog))+geom_boxplot()+facet_wrap(~variable)
```

---
## recasting data
```{r}
backtowide <- dcast(melteddata,formula = id+female+race+ses+schtyp+prog~variable)
```

---
## recasting data
```{r}
backtowide_byrace <- dcast(melteddata,formula = race~variable,fun.aggregate = mean)
backtowide_byrace
```
---
## recasting data
```{r}
backtowide_byraceses <- dcast(melteddata,formula = race+ses~variable,fun.aggregate = mean)
backtowide_byraceses
```

---
## Exercises 
1. datasets:EuStockMarkets is a dataset in wide format that lists the closing prices for 4 EU stock markets
  * Create a plot with 4 lines that tracks closing prices by date.
2. datasets::airquality is a dataset that tracks 4 variables of climate by month and day.
  * melt the data so that measurements are stacked
  * cast so that you get a mean monthly measurement by outcome measure.
 
---
## dplyr
```
install.packages("dplyr")
library(dplyr)
flights <- tbl_df(read.csv(url("http://cdn.rawgit.com/ylelkes/R_wav/master/data/flights.csv")))
```
```{r,echo=FALSE,warning=F,results=F,cache=TRUE}
library(dplyr)
flights <- tbl_df(read.csv(url("https://cdn.rawgit.com/ylelkes/R_wav/master/data/flights.csv")))
```

1. Most useful package (for me) in R. 
2. Another St. Wickham creation. 
3. A tool for data exploration and transformation

---
## 5 verbs
* filter: keep rows matching criteria
* select: pick columns by name
* arrange: reorder rows 
* mutate: add new variables
* summarise: reduce variables to values

--- 
## What is our general m.o. if wanted rows 1,3, & 4?
```{r}
df <- data.frame(   color = c("blue", "black", "blue", "blue", "black"),    value = 1:5)
head(df)
```

--- 
## What if we just wanted the "color" column

--- 
## What if we just wanted the "color" column
```{r}
select(df,color)
```

--- 
## Or if we didn't want the "color" column

--- 
## Or if we didn't want the "color" column
```{r}
select(df,-color)
```

--- 
## Take a look at the handout
```{r}
names(flights)
```
* What are some ways of selecting columns "dep_delay" and "arr_delay"

---
## Selecting _delay

```
select(flights, arr_delay, dep_delay)
select(flights, arr_delay:dep_delay)
select(flights, ends_with("delay"))
select(flights, contains("delay"))
```

---
## Arrange 

```{r}
arrange(df,value)
```

---
## Arrange 

```{r}
arrange(df,desc(value))
```

---
## Exercises 
* Order the flights by departure date and time.
* Which flights were most delayed?
* Which flights caught up the most time during the flight?

---
## Exercises
* arrange(flights, date, hour, minute) 
* arrange(flights, desc(dep_delay))
* arrange(flights, desc(arr_delay))
* arrange(flights, desc(dep_delay - arr_delay))


---
## Arrange 

```{r}
mutate(df,double= 2* value)
```

---
## Arrange 

```{r}
mutate(df,double= 2* value,quadruple=2*double)
```

---
## Exercises
* Compute speed in mph from time (in minutes) and distance (in miles).
* Which flight flew the fastest?
* Add a new variable that shows how much time was made up or lost in flight.

---
## Grouped Summaries
```{r}
summarise(df,total=sum(value))
```

---
## Grouped Summaries
```{r}
by_color <- group_by(df,color)
summarise(by_color,total=sum(value))
```

---
## Grouped summaries
* by_date <- group_by(flights, date)
* by_hour <- group_by(flights, date, hour)
* by_plane <- group_by(flights, plane)
* by_dest <- group_by(flights, dest) 

---
## Summary functions
* min(x), median(x), max(x), quantile(x, p)
* n(), n_distinct(), sum(x), mean(x) 
* sum(x > 10), mean(x > 10) 
* sd(x), var(x), iqr(x), mad(x)

---
## How might you summarise dep_delay for each day? 
```{r,echo=F}
library(ggplot2)
ggplot(flights,aes(dep_delay))+geom_histogram()
```

---
## Summarise by date

```
by_date <- group_by(flights, date)
delays <- summarise(by_date,   
mean = mean(dep_delay, na.rm = TRUE), 
median = median(dep_delay, na.rm = TRUE),
q75 = quantile(dep_delay, 0.75, na.rm = TRUE),
over_15 = mean(dep_delay > 15, na.rm = TRUE),
over_30 = mean(dep_delay > 30, na.rm = TRUE),
over_60 = mean(dep_delay > 60, na.rm = TRUE)  )
```

---
## Summarise by date
```
by_date <- group_by(flights, date)
no_missing <- filter(flights, !is.na(dep))
delays <- summarise(by_date,   
mean = mean(dep_delay, na.rm = TRUE), 
median = median(dep_delay, na.rm = TRUE),
q75 = quantile(dep_delay, 0.75, na.rm = TRUE),
over_15 = mean(dep_delay > 15, na.rm = TRUE),
over_30 = mean(dep_delay > 30, na.rm = TRUE),
over_60 = mean(dep_delay > 60, na.rm = TRUE)  )
```