---
title: "Make these plots"
author: "ylelkes"
date: "September 7, 2015"
output: pdf_document
---

## Make these plots

Use the Fair data in the Ecdat package
[Read about it here](https://cran.r-project.org/web/packages/Ecdat/Ecdat.pdf)

## Part 1
```{r,echo=F,warning=F,message=F}
library(ggplot2)
library(Ecdat)
ggplot(Star,aes(tmathssk))+geom_histogram()
ggplot(Star,aes(tmathssk))+geom_density()
ggplot(Star,aes(tmathssk,fill=sex))+geom_density()
ggplot(Star,aes(tmathssk,fill=sex))+geom_density(alpha=.5)
ggplot(Star,aes(x=sex,y=tmathssk,fill=sex))+geom_boxplot(colour="purple")
ggplot(Star,aes(x=totexpk,y=tmathssk))+geom_smooth()
ggplot(Star,aes(x=totexpk,y=tmathssk))+geom_smooth()+geom_point()
```

## Part 2
```{r,echo=F,warning=F,message=F}
ggplot(Star,aes(x=tmathssk,y=treadssk,colour=race))+geom_point()+facet_wrap(~sex)
ggplot(Star,aes(x=tmathssk,y=treadssk))+geom_point()+facet_grid(race~classk)
ggplot(Star,aes(x=tmathssk,y=treadssk,colour=sex))+geom_point()+facet_grid(race~classk)+geom_smooth(method=lm,colour="black")
ggplot(Star,aes(x=tmathssk,y=treadssk,shape=sex,colour=freelunk))+geom_point()+facet_grid(race~classk)+geom_smooth(method=lm)
```

Still have time? Change the axis labels, add a title

## Part 3

The file [climatetemperatures.RData](https://github.com/ylelkes/R_wav/blob/master/data%20examples/climatetemperatures.RData?raw=true) is a list with two data frames. Temperature_by_year is used in the first plot; futureeuropetemp is based off of 15 or so climate models predicting the temperature of european countries in the end of the 21st centure. Hint: The 95 percent confidence intervals (which are plotted below are 1.96 times the standard error over/under the mean)

```{r,echo=F,warning=F,message=F}
load("../data examples/climatetemperatures.RData")
ggplot(temperature_by_year,aes(x=year,ymin=min,ymax=max,y=temperature))+geom_ribbon(alpha=.5,fill="red")+geom_line()
ggplot(futurepast,aes(x=temperature,xmin=temperature-1.96*se,xmax=temperature+1.96*se,y=locator,colour=time))+geom_point()+geom_errorbarh()
```


Have more time? 
Change the theme. Start exploring scales, and change the colours of the points. Change the y-axis to go from the minimum to the maximum. Can you change the title of the legend? 