---
title: "Problem Set 1 Answers"
author: "ylelkes"
date: "September 14, 2015"
output: html_document
---
For now, copy and paste your commands and results (including figures) into a word document.

1. Suppose you track your commute times for two weeks (10 days) and you find the following times in minutes: 17 16 20 24 22 15 21 15 17 22

```{r}
commute <- c(17, 16,20,24,22,15,21,15,17,22)
```  
  
  
	* Use the function max to find the longest commute time, the function mean to find the average and the function min to find the minimum.

```{r}
max(commute)
mean(commute)
min(commute)
```  

* Oops, the 24 was a mistake. It should have been 18. How can you fix this without rewriting the vector? Do so, and then find the new average.
```{r}
commute[4]=18
commute[commute==24]=18
commute <- car::recode(commute,"24=18")
mean(commute)
```  


* How many times was your commute 20 minutes or more? Hint: try using sum() and a logical statement. [Read here about logical operators](http://www.statmethods.net/management/operators.html)

```{r}
commute>20
sum(commute>20)
```

* What percent of your commutes are less than 17 minutes? 
```{r}
commute>20
sum(commute<17)/length(commute)*100
scales::percent(sum(commute<17)/length(commute))
```

* Replace the 5th element of the vector with NA. Calculate the new mean. 

```{r}
commute[5]<- NA
mean(commute)
mean(commute,na.rm=T)
```

2. The Crimean War
	
	* Install the HistData package and load Nightingale. Read about it [here](https://vincentarelbundock.github.io/Rdatasets/doc/HistData/Nightingale.html)
	* How many people died of Wounds? How many people died of Disease?
```{r}
library(HistData)
data(Nightingale)
sum(Nightingale$Wounds)
sum(Nightingale$Disease)
```
	* Create a piechart using the pie() command. You should have a slice for # of ppl dying of wounds, # of ppl dying of disease, # of ppl dying of other. Make sure to label all the slices. Give your graph a title. If you wish, change the colors. 

```{r}
pie(x = c(sum(Nightingale$Wounds),sum(Nightingale$Disease),sum(Nightingale$Other)),labels = c("Wounds","Disease","Other"),main = "Causes of Death in the Crimean War",col = c("red","blue","green"))

with(Nightingale, pie(x = c(sum(Wounds),sum(Disease),sum(Other)),labels = c("Wounds","Disease","Other"),main = "Causes of Death in the Crimean War",col = c("red","blue","green")))
```


3. Loading data into R
	* We didn't get to this in class, but loading data into R will be central to your life as a data analysist. 
	* Download the following files into your working directory (hint: use getwd())
		* If you want to make it more challenging, read it directly from the web (without downloading it to your working directory).  
		* https://raw.githubusercontent.com/ylelkes/R_wav/master/data%20examples/Countries-Europe.csv
		* https://github.com/ylelkes/R_wav/blob/master/data%20examples/child_data.sav?raw=true
		* https://github.com/ylelkes/R_wav/blob/master/data%20examples/GaltonFamilies.dta?raw=true
    
```{r}
library(RCurl)
europe <- rio::import(("https://rawgit.com/ylelkes/R_wav/master/data%20examples/Countries-Europe.csv"))
childdata <- rio::import("https://rawgit.com/ylelkes/R_wav/master/data%20examples/child_data.sav")
galton <- rio::import("https://rawgit.com/ylelkes/R_wav/master/data%20examples/GaltonFamilies.dta")
```    
	* Using read.csv, the foreign package, or haven load each one of these datasets into R.
	* For: "Countries-Europe.csv" (let's call that object europe)
		* What is the median population of Europe?
```{r}
median(europe$population)
```
* What is the mean population/land area

```{r}
mean(europe$population/europe[,7])
mean(europe$population/europe$land)
```


* If you replace X and Y in the following, you will get most populated country in Europe:
			``` 
			europe$X[europe$Y==max(europe$population)]
			```
```{r}
europe$name[europe$population==max(europe$population)]
```


* What variable is X? What variable is Y?
		* In your own words, describe what this command is doing..
	
  
  * For GaltonFamilies.dta:
		* using the cor() command, what is the correlation between a child's height and his mother's height and what is the correlation between the child's height and the father's height? 

```{r}
fatherchild <- cor(galton$childHeight,galton$father)
motherchild <- cor(galton$childHeight,galton$mother)
```
		* Use a logical statement, get R to confirm that the first correlation does not equal the second correlation.
```{r}

fatherchild == motherchild
```

* Using lm(), is there a relationship between the father's height and the number of children he has?
	
```{r}
summary(lm(father~childHeight,galton))
```

* For child_data.sav
		* What is the memory span of child with the highest IQ?

```{r}

childdata$MEM_SPAN[childdata$IQ==max(childdata$IQ)]
```


* Create a correlation table for the entire dataset

```{r}
cortab <- cor(childdata)
```

		* From the correlation table, extract the correlation between memory span & IQ, the correlation between age and reading ability, and the correlation between IQ and reading ability into a vector

```{r}
c(cortab[2,3],cortab[1,4],cortab[3,4])
```

* write that vector to a csv file.
```{r}
write.csv(c(cortab[2,3],cortab[1,4],cortab[3,4]),file="cortabl.csv")
```

	





